---
title: "Intro to scRNAseq"
output: html_document
---

```{r libraries, message=FALSE, warning=FALSE}

library(Seurat)
library(ggplot2)

```

#cell conts and metadata from local dir
```{r read in data}

getwd() #check local working dir

xim_meta <- read.csv("metadata_file_1.txt", header = TRUE, sep = "\t")

xim_counts <- read.table("counts_file_1.txt", header = TRUE, sep = "\t", check.names = FALSE)

all(row.names(xim_meta) == colnames(xim_counts)) #are all the cells the same?

```

#create seurat objects from single file
```{r seurat object}

xim_seurat<-CreateSeuratObject(counts = xim_counts, meta.data = xim_meta)

xim_seurat_filter_first<-CreateSeuratObject(counts = xim_counts, meta.data = xim_meta, min.cells = 10)

```

#Filter lowly expressed genes
```{r lowly expressed genes}

#look at raw data
counts <- rowSums(xim_seurat@assays$RNA@counts != 0)
counts <- data.frame(counts=counts, row.names = rownames(xim_seurat))
hist(counts$counts, col = 'red', breaks = 50, main ="feature/cell counts, non filter")

#filter min number of cells expressing any gene
counts_filter_now <- data.frame(counts = counts[counts$counts >= 10,, drop=FALSE]) #set minimum cells 
hist(counts_filter_now$counts, col = 'skyblue3', breaks = 50, main ="feature/cell counts, filter now")

#look at data filtered via create seurat object
counts_filter_first <- rowSums(xim_seurat_filter_first@assays$RNA@counts != 0)
counts_filter_first<-data.frame(counts = counts_filter_first, row.names = rownames(xim_seurat_filter_first))
hist(counts_filter_first$counts, col = 'green', breaks = 50, main ="feature/cell counts, seurat filter first")

#how any genes did we remove?
paste("number of removed genes:", nrow(counts)-nrow(counts_filter_now))
paste("percent of removed genes:", (nrow(counts)-nrow(counts_filter_now))/nrow(counts)*100)

```

#remove lowly expressed transcripts from raw seurat object, using "filter now" geens
```{r}
gene.keep<-rownames(counts_filter_now)
xim_seurat<-xim_seurat[rownames(xim_seurat) %in% c(gene.keep),]
```

#Quality Control
```{r QC}

#Calc %Mito Genes
xim_seurat[["percent.mt"]] <- PercentageFeatureSet(xim_seurat, pattern = "^mt-")
xim_seurat[["percent.ribo"]]<-PercentageFeatureSet(xim_seurat, pattern = "^Rp")

#Visualize QC metrics as a violin plot
VlnPlot(xim_seurat,
        features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.ribo"),
        ncol = 4)


FeatureScatter(xim_seurat, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")+
  NoLegend() 
FeatureScatter(xim_seurat, feature1 = "nFeature_RNA", feature2 = "percent.mt")+
  NoLegend() 

```

#Filter
```{r Filter}
xim_seurat <- subset(xim_seurat, subset = nFeature_RNA > 300 & nFeature_RNA <= 3000 & 
                       percent.mt <= 10 & percent.ribo <=13) 


FeatureScatter(xim_seurat, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")+
  NoLegend() 
FeatureScatter(xim_seurat, feature1 = "nFeature_RNA", feature2 = "percent.mt")+
  NoLegend() 

```

#Normalize, scale, findvariable features, and plot - the Seurat pipeline
```{r Norm and scale}
xim_seurat <- NormalizeData(xim_seurat) 
xim_seurat <- FindVariableFeatures(xim_seurat, selection.method = "vst", nfeatures = 2000)
xim_seurat <- ScaleData(xim_seurat)

```

#Example of log10 normalization
```{r log10 example}
hist(as.matrix(xim_seurat@assays$RNA@counts),
     main = "Raw Counts")
hist(as.matrix(xim_seurat@assays$RNA@data),
     main = "Seurat norm and scaled") 
hist(log10(as.matrix(xim_seurat@assays$RNA@counts)+1),
     main = "Raw Counts log10+1")
```

```{r dimensional reduction}
xim_seurat <- RunPCA(xim_seurat,
                     npcs = 50,
                     features = VariableFeatures(object = xim_seurat))
xim_seurat <- RunUMAP(xim_seurat,
                      dims = 1:30,
                      n.neighbors = 30,
                      min.dist = 0.3)

DimPlot(xim_seurat, reduction = "umap") + NoLegend()

#if you want to try a tSNE
#xim_seurat <- RunTSNE(xim_seurat, dims = 1:30)
#DimPlot(xim_seurat, reduction = "tsne") + NoLegend()

```

#View dataset annotations
```{r data annotations}

plot <- DimPlot(xim_seurat,
        group.by = "cell_type_age",
        label = T,
        repel = T)

plot + NoLegend()

```

#Finding cluster markers
```{r}

```

